Statically Typed String Sanitation Inside a Python

### 2014-08-21 sanitation-psp14/sanitation-psp14-short.tex
JA sec 2

The $\lambdas$ language supports two forms of coercion. Safe coercion only allows coercsion between 
strings types which are guaranteed to be safe. This form of coercion is useful because the replacement operation (introduced below in Section~\ref{sec:replace}) is often more conservative than absolutely necessary.
Conversely, unsafe coercion -- which we refer to as a checked coercion -- 
allows for potential unsafe type casts. Our semantics for check ensures that only safe values are used. Whenever a value is determined to be unsafe at runtime, a default value $e_2$ is used instead.

Summarily, 
 \textsc{S-T-SafeCoerce} allows only safe coercions between string types by
    expoiting the decidability of language inclusion, while
\textsc{S-T-Check} allows casts between strings that cannot be guaranteed at compile time, but inserts a runtime check a runtime check.

@@

The premier operation for manipulating strings in $\lambdas$ is string substitution.
On strings, the string substitution operator is familiar, and roughly analogous to \lstinline{str_replace} in PHP and \lstinline{String.replace} in Java.
The definition of \textsc{S-T-Case} requires a definition of replacement for regular expressions (or languages) corresponding to string replacement.
In this section, we define this operation.

Both string and language replacement are defined extra-linguistically.
The system $\lambdas$ is defined in terms of these functions.
The function $\lsubst{r}{s_1}{s_2}$ replaces every substring of $s_1$ matching $r$ with $s_2$.
In an analogous manner, the function $\lreplace{r}{r_1}{r_2}$ replaces every sublanguages of $r_1$ matching $r$ with $r_2$.
If an intuiition for this operation is not clear, it may be helpful to think in terms replacing sub-automata.

@@

\begin{defn}[$\tt{lreplace}$]
  The relation $\lreplace{r}{r_1}{r_2} = r'$ relates $r, r_1,$ and $r_2$ to
  a language $r'$ containing all strings of $r_1$ except that any substring $s_{pre} s s_{post} \in \lang{r_1}$ where $s \in \lang{r}$
  is replaced by the set of strings $s_{pre} s_2 s_{post}$ for all $s_2 \in \lang{r_2}$ (the prefix and postfix positions may be empty).
\end{defn}

@@

\begin{lem}
  If $\sctx \vdash e : \stringin{r}$ then $r$ is a well-formed regular expression.
\end{lem}
\begin{proof}
  The only non-trivial case is S-T-Replace, which follows from lemma \ref{thm:total}. 
\end{proof}
\begin{lem}
If $\tctx \vdash \iota : \regex$ then $\iota \treduces \rx{r}$ such that $r$ is a well-formed regular expression. 
\end{lem}

@@

\begin{thm}[Type Safety.] \label{thm:scorrect}
  If $\sctx \vdash e : \sigma$ 
  then $e \sreduces e'$ and $\sctx \vdash e' : \stringin{r}$.
\end{thm}
\begin{proof}
By induction on the typing relation.
The S-T-Concat case requires Lemma \ref{thm:regexprops} 
and the S-T-Replace case appeals to Lemma \ref{thm:substcorrespondence}.
\end{proof}

@@

>>>>

The $\lambdas$ language supports two forms of coercion. Safe coercion only allows coercion between 
strings types which are guaranteed to be safe. This form of coercion is useful because the replacement operation (introduced below in Section~\ref{sec:replace}) is often more conservative than absolutely necessary.
Conversely, unsafe coercion -- which we refer to as a checked coercion -- 
allows for potential unsafe type casts. Our semantics for check ensures that only safe values are used. Whenever a value is determined to be unsafe at runtime, a default value $e_2$ is used instead.

Summarily, 
 \textsc{S-T-SafeCoerce} allows only safe coercions between string types by
    expoiting the decidability of language inclusion, while
\textsc{S-T-Check} allows casts between strings that cannot be guaranteed at compile time, but inserts a runtime check.

An alternative to coercion is subtyping for constrained string types. For simplicity, we do not include this relation; however, subtyping for constrained strings is present in \cite{fulton13} and that treatment extends to this system.

@@

The premier operation for manipulating strings in $\lambdas$ is string substitution.
On strings, the string substitution operator is familiar, and roughly analogous to \lstinline{str_replace} in PHP and \lstinline{String.replace} in Java.
The definition of \textsc{S-T-Replace} requires a definition of replacement for regular expressions (or languages) corresponding to string replacement.
In this section, we define this operation.

Both string and language replacement are defined extra-linguistically.
The system $\lambdas$ is defined in terms of these functions.
The function $\lsubst{r}{s_1}{s_2}$ replaces every substring of $s_1$ matching $r$ with $s_2$.
In an analogous manner, the function $\lreplace{r}{r_1}{r_2}$ replaces every sublanguage of $r_1$ matching $r$ with $r_2$.
If the intuition for this operation is not clear, it may be helpful to think in terms of replacing sub-automata.

@@

\begin{defn}[$\tt{lreplace}$]
  The relation $\lreplace{r}{r_1}{r_2} = r'$ relates $r, r_1,$ and $r_2$ to
  a language $r'$ containing all strings of $r_1$ except that any substring $s_{pre} s s_{post} \in \lang{r_1}$ where $s \in \lang{r}$
  is replaced by the set of strings $s_{pre} s_2 s_{post}$ for all $s_2 \in \lang{r_2}$ (the prefix and postfix positions may be empty). This procedure saturates the string.
\end{defn}

@@

\begin{lem}
  If $\sctx \vdash e : \stringin{r}$ then $r$ is a well-formed regular expression.
\end{lem}
\begin{proof}
  The only non-trivial case is S-T-Replace, which follows from lemma \ref{thm:total}. 
\end{proof}

@@

\begin{thm}[Type Safety.] \label{thm:scorrect}
  If $\sctx \vdash e : \sigma$ 
  then $e \sreduces v$ and $\sctx \vdash v : \sigma$.
\end{thm}
\begin{proof}
By induction on the typing relation.
The S-T-Concat case requires Lemma \ref{thm:regexprops} 
and the S-T-Replace case appeals to Lemma \ref{thm:substcorrespondence}.
\end{proof}

@@

\begin{lem}
If $\tctx \vdash \iota : \regex$ then $\iota \treduces \rx{r}$ such that $r$ is a well-formed regular expression. 
\end{lem}
